---
title: "assignment.Rmd"
author: "Nandakishor Koka"
date: "June 22, 2014"
output: html_document
---

Initial Setup - Load the Caret package; Make sure all dependencies are installed  a

```{r}
library(caret) 
set.seed(9) 
```
Load the data sets 

```{r}
hlmtrain <- read.csv("data/pml-training.csv", header=TRUE)
hlmtest <- read.csv("data/pml-testing.csv", header=TRUE)
```

Partition the training data for training and cross validation 
```{r} 
inTrain <- createDataPartition (y=hlmtrain$classe, p = 0.80, list=FALSE )
training <- hlmtrain[inTrain,]
testing <- hlmtrain[-inTrain,]
```

Remove variables that have near zero variance 

```{r}
nzv <- nearZeroVar(hlmtrain)
training <- training[, -nzv] 
testing <- testing[, -nzv]
```

Filter for only numerical columns 
```{r}
nfs = which(lapply(training,class) %in% c('numeric'))
preproc <- preProcess(training[,nfs], method=c('bagImpute'))
training_new <- cbind(training$classe, predict(preproc, training[,nfs]))
testing_new <- cbind(testing$classe, predict(preproc, testing[,nfs]))
```

Correct the label 
```{r} 
names(training_new)[1] <- "classe"
names(testing_new)[1] <- "classe"
```

Check for variable with high correlation and remove these 
```{r} 
nfs2 = which(lapply(training_new,class) %in% c('numeric'))
dcor <- cor(training_new[,nfs2])
highcor <- findCorrelation(dcor, cutoff = 0.90)
training_new <- training_new[,-highcor]
``` 

Train the model. Using Gradient boosting as it provides better flexibility in tweaking the model. Limiting the number to allow processing on small machines.  
```{r] 
#fitControl <- trainControl(method = "repeatedcv",number = 10,repeats = 10)
fitControl <- trainControl(number=1)
model = train(classe ~ ., data = training_new, method = "gbm",  verbose = TRUE, trControl = fitControl)
```

In-Sample accuracy: 
```{r} 
training_new_pred <- predict(model, training_new) 
print(confusionMatrix(training_new_pred, training_new$classe))
``` 

Out-of-Sample accuracy: 
```{r} 
testing_new_pred <- predict(model, testing_new) 
print(confusionMatrix(testing_new_pred, testing_new$classe))
```

Final Test: 
```{r} 
hlmtest <- hlmtest[, -nzv]
final_test <- predict(preproc, hlmtest[,nfs])
final_res <- predict(model, final_test)
print (final_res ) 
```


